---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

```{r echo=FALSE}
# install.packages("ggplot2")
# install.packages("lubridate")
# install.packages("xtable")
library(ggplot2)
library(lubridate)
library(xtable)
# library(tinytex)
# install_tinytex()
```

## Loading and preprocessing the data
Convert the date from char to date
```{r}
rawdata <- read.csv("activity.csv")
data <- rawdata
data$date <- as.Date(data$date)
```

## What is mean total number of steps taken per day?
This can be solved with the aggregate functions:
```{r}
sumVec <- with(data,aggregate(steps,list(date),sum,na.rm=TRUE))
colnames(sumVec) <- c("Date", "Sum")
meanVec <- with(data,aggregate(steps,list(date),mean,na.rm=TRUE))
colnames(meanVec) <- c("Date", "Mean")
medianVec <- with(data,aggregate(steps,list(date),median,na.rm=TRUE))
colnames(medianVec) <- c("Date", "Median")
hist(sumVec$Sum, main = "Histogram of total daily steps", xlab = "Number of steps")

meanmed <- merge(meanVec,medianVec)
meanmed
```

## What is the average daily activity pattern?

```{r}
meanVec2 <- with(data,tapply(steps,interval,mean,na.rm=TRUE))
plot(unique(data$interval),meanVec2,type="l", main ="Average daily activity pattern", xlab = "Interval", ylab = "Number of steps")
maxint <- rownames(meanVec2)[which(meanVec2 == max(meanVec2))]
```
The `r maxint` interval contains the maximum number of steps, on average across all the days in the dataset.

## Imputing missing values
```{r}
naIdx = which(is.na(data))
z <- length(naIdx)
df <- data.frame(meanVec2)
newData <- data
```
There are `r z` NA values in the dataset.
The strategy to deal with the NAs is to plug in the calculated mean for the corresponding 5-minute interval.
```{r}
for (i in 1:length(naIdx)) {
  int <- data$interval[naIdx[i]]
  newSteps <- df[as.character(int),"meanVec2"]
  newData$steps[naIdx[i]] <- newSteps
}

sumVec <- with(newData,tapply(steps,date,sum,na.rm=TRUE))
sumVec <- data.frame(key=names(sumVec), value=sumVec)
# p <- ggplot(data=sumVec, aes(x=key, y=value)) +
# +     geom_bar(stat="identity")

meanVec <- with(newData,aggregate(steps,list(date),mean,na.rm=TRUE))
colnames(meanVec) <- c("Date", "Mean")
medianVec <- with(newData,aggregate(steps,list(date),median,na.rm=TRUE))
colnames(medianVec) <- c("Date", "Median")
meanmed <- merge(meanVec,medianVec)
meanmed
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
newData$wd <- wday(newData$date, week_start = 1)
newData$wd[which(newData$wd < 6)]="weekday"
newData$wd[which(newData$wd != "weekday")]="weekend"
newData$wd <- factor(newData$wd)

weekdayMean <- with(newData,aggregate(steps, by=list(interval,wd), FUN=mean, na.rm=TRUE))

s <- ggplot(weekdayMean, aes(Group.1, x)) + geom_line()
s <- s + facet_grid(rows=vars(Group.2))
s <- s + labs(x = "Interval", y = "Number of steps")
s
```